### 2017-04-04 sec-intro.tex
Update paper to incorporate review A and review B.
Also update our response to these two reviewers to be more concrete
about the exact changes we have made to the paper and where to find
them.

Our work is motivated by the following observations:
\begin{enumerate}
  \item Student behavior is complicated and cannot necessarily be captured
      sufficiently by rule-based methods such as those explored by
      \citet{Kizilcec:2013:LAK} and \citet{Davis:2016:EDM}. We instead
      propose to treat student behavior patterns as being characterized
      (represented) via a sequence of \emph{latent states}. This allows us
      to automatically capture patterns that we might not have been able to
      articulate clearly a priori via a series of rules, and also allows us
      to model the inherent uncertainty in assigning a student's behavior
      to a particular pattern or group.
  \item Student behavior can vary over time. Previous models that treat students
      as exhibiting only one behavioral pattern over
      time~\cite{Faucon:2016:EDM} miss out on the opportunity to understand
      student behavior dynamics in a course. We propose a latent space
      model with {\em latent state transitions} to flexibly model the
      dynamics.
  \item Analysis of student behavior can and should be performed at varying
      levels of granularity. This requires us to aggregate data over time
      with \emph{different levels of resolution}; existing models tend to come
      with a particular assumption about the resolution of time they
      consider~\cite{Faucon:2016:EDM, Kizilcec:2013:LAK, Shih:2010:EDM}. We
      propose a more flexible model to accommodate different levels of
      resolution.
\end{enumerate}

Thus, what we propose is a \emph{latent variable approach} to mining student behavior
patterns that is \emph{probabilistic} for inference and 
\emph{flexible to model state changes over different time resolutions}. 
More specifically, we propose a novel two-layer hidden Markov model (2L-HMM) to
 discover latent student behavior patterns via
unsupervised learning on large collections of student behavior observation
sequences. Evaluation results on a MOOC data set on Coursera demonstrate that 2L-HMM can effectively 
discover a variety of interesting interpretable student behavior patterns at different levels of resolution, many of which are beyond what existing approaches can discover. 2L-HMM further enables easy use of the discovered patterns for discriminative analysis such as prediction of learning outcome. 
Since our proposed methods are unsupervised, they can potentially be applied to any MOOC data without requiring any manual work to facilitate understanding of student behaviors and their variations, opening up many possibilities for developing intelligent tutoring systems that can adaptive to student behavior.

>>>>

We view our model as a component of a system that enables collaboration
between the machine and a human instructor to extract knowledge from these
large collections of MOOC data. Automatically extracting interpretable
patterns from the clickstream data associated with MOOCs is a necessary
step in order for instructors to identify the hidden knowledge in these
massive interaction datasets. Without the availability of a suitable model
for identifying these behavioral patterns, instructors are not empowered to
use this available data to improve their courses without expending
extraordinary amounts of manual effort.

Our proposed model is motivated by the following observations:
\begin{enumerate}
  \item Student behavior is complicated and cannot necessarily be captured
      sufficiently by rule-based methods such as those explored by
      \citet{Kizilcec:2013:LAK} and \citet{Davis:2016:EDM}. We instead
      propose to treat student behavior patterns as being characterized
      (represented) via a sequence of \emph{latent states}. This allows us
      to automatically capture patterns that we might not have been able to
      articulate clearly a priori via a series of rules, and also allows us
      to model the inherent uncertainty in assigning a student's behavior
      to a particular pattern or group.
  \item Student behavior can vary over time. Previous models that treat students
      as exhibiting only one behavioral pattern over
      time~\cite{Faucon:2016:EDM} miss out on the opportunity to understand
      student behavior dynamics in a course. We propose a latent space
      model with {\em latent state transitions} to flexibly model the
      dynamics.
  \item Analysis of student behavior can and should be performed at varying
      levels of granularity. This requires us to aggregate data over time
      with \emph{different levels of resolution}; existing models tend to come
      with a particular assumption about the resolution of time they
      consider~\cite{Faucon:2016:EDM, Kizilcec:2013:LAK, Shih:2010:EDM}. We
      propose a model that is agnostic to the time resolution considered,
      allowing it to be applied at different levels of resolution more
      naturally.
\end{enumerate}

Thus, what we propose is a \emph{latent variable approach} to mining student behavior
patterns that is \emph{probabilistic} for inference and 
does \emph{not} force assumptions about time resolutions, making it
\emph{flexible to model state changes over different time resolutions} more
easily.  More specifically, we propose a novel two-layer hidden Markov
model (2L-HMM) to discover latent student behavior patterns via
unsupervised learning on large collections of student behavior observation
sequences.  Evaluation results on a MOOC data set on Coursera demonstrate
that the 2L-HMM can effectively discover a variety of interesting interpretable
student behavior patterns at different levels of resolution, many of which
are beyond what existing approaches can discover. We show that the patterns
uncovered by the 2L-HMM capture meaningful behavior by quantitatively
showing that features extracted from a trained 2L-HMM correlate with
learning outcomes. Since our proposed methods are
unsupervised, they can potentially be applied to any MOOC data without
requiring manual annotation effort at the level of sequences. Instead,
instructors are empowered to use the latent patterns that the 2L-HMM can
discover from raw data to extract knowledge about the behaviors his/her
students exhibit in the MOOC.
